---
title: 'Computational Musicology: David Bowie'
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    theme: flatly
    self_contained: false
editor_options:
  chunk_output_type: console
---
<style>
.navbar-brand {
  color:black!important;
}
</style>
```{r setup, include=FALSE}
source('bowie.R')
```


### Introduction

David Bowie was a musical chameleon. What patterns can be found in his many-colored oeuvre?

This project uses the R packages [Spotifyr](https://www.rcharlie.com/spotifyr/) and [compmus](https://jaburgoyne.github.io/compmus/) to process data provided by the [Spotify Web API](https://developer.spotify.com/documentation/web-api/).

The corpus consists of the `r length(unique(bowie$album))` studio albums. Compilations, live albums and rarities are excluded from the selection, to prevent overlap.

The analysis focuses on the **Berlin Trilogy**: the albums Low (1977), Heroes (1977), and Lodger (1979).

***

Playlist for the corpus (`r nrow(bowie)` tracks)
<br><br>
<iframe style="border-radius:12px" src="https://open.spotify.com/embed/playlist/4L34OFuKhWwCB9akiVC2Ak?utm_source=generator&theme=0" width="100%" height="380" frameBorder="0" allowfullscreen="" allow="autoplay; clipboard-write; encrypted-media; fullscreen; picture-in-picture"></iframe>


### Instrumentation

```{r}
p <- bowie_scatterplot(x_aes = 'acousticness',
                       y_aes = 'instrumentalness')

ggplotly(p, tooltip = c('text'))
```

***

Brian Eno was an important collaborator on the Berlin Trilogy. His influence is especially apparent in three consecutive tracks near the end of the album "Heroes": Sense of Doubt, Moss Garden, and Neuköln. These tracks form a cohesive whole and have been called the **Trilogy Within the Berlin Trilogy**, visualized here by the triangle. The style of these acoustic-like *ambient* instrumentals is vintage Eno.

**Acousticness** predicts the absence of electrical amplification of musical instruments. Although the Trilogy Within sounds more acoustic than most of the other tracks, there is no clear separation in this respect.

**Instrumentalness** predicts whether a track is without vocals. Two albums in the Berlin Trilogy contain many instrumental tracks: "Heroes" and Low. Most tracks in the corpus feature vocals, resulting in a mean instrumentalness of `r format(mean(bowie$instrumentalness), digits = 3, scientific=FALSE, drop0trailing=TRUE)` and a median of `r format(median(bowie$instrumentalness), digits = 3, scientific=FALSE, drop0trailing=TRUE)`.


### Emotion

```{r}
x_dim <- 'valence'
y_dim <- 'energy'
highlight_tracks <- c('The Secret Life of Arabia')
emo_color <- 'grey30'
p <- bowie_scatterplot(x_aes = x_dim,
                       y_aes = y_dim) +
  geom_point(data = bowie[bowie$track %in% highlight_tracks,],
             pch = 21, fill = NA, size = 3, stroke = 0.2) +
  annotate(geom = 'text', colour = emo_color,
           label = c('ANGRY','HAPPY','SAD','CALM'),
           x = c(0.08,0.93,0.05,0.94), y = c(0.96,0.96,0.03,0.03))

ggplotly(p, tooltip = c('text')) %>%
  # Disable tooltips for emotion labels
  style(hoverinfo = 'none', traces = c(10))
```

***

**Valence** predicts the musical positiveness of a track, and ranges from negative to positive affect. **Energy** is a perceptual measure of intensity and activity. Energy is roughly analogous to the concept of arousal in psychology. The valence-arousal chart is a popular map of the four basic emotions.

Bowie's oeuvre covers a large part of the map. Like most music, the tracks are not very calm. The tracks also shun the angry corner, where some of the darker alternative genres live. The Trilogy Within has a somewhat gloomy feel, and resides near the sad corner. The next and last track on the album "Heroes", The Secret Life of Arabia, emerges as one of the happiest Bowie tracks. It is accentuated on the map. The Berlin Trilogy in general doesn't stand out in comparison to the other albums.


### Chroma

```{r}
track_name <- 'Sense of Doubt'
track_uri <- gsub('spotify:track:', '',
                    bowie[bowie$track == track_name,]$track.uri)

data <-
  compmus::get_tidy_audio_analysis(track_uri) %>%
  select(segments) %>% tidyr::unnest(segments) %>%
  select(start, duration, pitches)

chromagram <- data %>%
  mutate(pitches = purrr::map(pitches, compmus_normalise, 'euclidean')) %>%
  compmus_gather_chroma %>%
  ggplot(aes(x = start + duration / 2,
             y = pitch_class,
             width = duration,
             fill = value,
             text = paste0(round(value, digits = 2)))) +
  geom_tile() +
  labs(x = 'Time (s)', y = NULL, fill = 'Magnitude') +
  theme_minimal() +
  theme(plot.background = element_rect(fill = bg_color),
        panel.background = element_rect(fill = bg_color)) +
  scale_fill_gradient(low = 'black', high = 'white')

ggplotly(chromagram, tooltip = c('text'))
```

***

Sense of Doubt is the first of three instrumental tracks on the album "Heroes", the aforementioned Trilogy Within. The previous plot categorized these instrumentals as sad music. In fact, `r bowie[bowie$valence == min(bowie$valence),]$track` is the track with lowest valence in the corpus: `r min(bowie$valence)`.

In the chromagram we see the pitchless white noise of the first seven seconds. The pitch class A is very present throughout the track. There is a descending four-note motif, C-B-Bb-A, where the A is held. For the key, Spotify returns pitch class `r (key <- bowie[bowie$track == 'Sense of Doubt',]$key)`. This corresponds to row `r c('C', 'C#|Db', 'D', 'D#|Eb', 'E', 'F', 'F#|Gb', 'G', 'G#|Ab', 'A', 'A#|Bb', 'B')[1+key]` in the chromagram.


### Self-Similarity

```{r}
ssm <-
  get_tidy_audio_analysis(track_uri) %>%
  compmus_align(bars, segments) %>%
  select(bars) %>% tidyr::unnest(bars) %>%
  mutate(pitches = map(segments, compmus_summarise, pitches,
                       method = 'rms', norm = 'euclidean')) %>%
  mutate(timbre = map(segments, compmus_summarise, timbre,
                      method = 'mean'))
ssm %>%
  compmus_self_similarity(timbre, 'cosine') %>%
  ggplot(aes(x = xstart + xduration / 2,
             y = ystart + yduration / 2,
             width = xduration,
             height = yduration,
             fill = d)) +
  geom_tile() +
  coord_fixed() +
  # scale_fill_viridis_c(option = 'E', guide = 'none') +
  scale_fill_gradient(low = 'black', high = 'white') +
  theme_classic() +
  labs(x = '', y = '') +
  theme(legend.position='none')
```

***

In Sense of Doubt, the low-register four-note piano riff is juxtaposed with a high pitched synth line. The self-similarity matrix visualizes similarities and differences between time segments.

<iframe style="border-radius:12px" src="https://open.spotify.com/embed/track/1d89lSLqTpMMpcKBWzVWcm?utm_source=generator" width="100%" height="380" frameBorder="0" allowfullscreen="" allow="autoplay; clipboard-write; encrypted-media; fullscreen; picture-in-picture"></iframe>
